### 2016-03-01 ch02-qlogms/excitation-design.tex
Vuistregels etc.

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

@@

  \begin{definition}
  For the implementation of a quasi-logarithmic frequency grid,  the approximate equality \eqref{eq:excitation:qlogGrid} does not define the grid unambiguously.
  The quasi-logarithmic grids depicted in this chapter are constructed using the relationship
  \begin{align}
       \label{eq:excitation:qlog:gridDef}
    f_k & = \alpha_k(N_k) f_{k-1}  
        \quad
            \forall k \in \set{1,\ldots,F}
\end{align}
with
\begin{align}
     \label{eq:excitation:qlog:roundAlpha}
      \alpha_k(N_k) & \isdef 
          \round{   \frac{\alpha^{N_k} f_{k-1}}{\Delta f}}    
           \frac{\Delta f}{f_{k-1}}
      \approx \alpha \text{, and}\\
         \label{eq:excitation:qlog:alphaPowerDef}
          N_k & \isdef 
            \min 
            \set{n | n \in \NaturalNumbers, \alpha_k(n) > 1}
     \text{.}
  \end{align}
  \end{definition}
  In the equation above, $\round{\placeholder}$ denotes rounding towards the nearest integer according to the \IEEEfloat standards and as implemented in the \MATLAB function \mcode{round}.
  This satisfies the approximate definition.

  \begin{remark}
  The use of different rounding strategies --- e.g. the \mcode{ceil} function $\ceil{}$  and \mcode{floor} function $\floor{}$ --- in \eqref{eq:excitation:qlog:roundAlpha}, allows for slightly different guarantees on the effective $\alpha_k$ attained by the grid.
  By denoting $\alpha_k^{\placeholder}$ as the effective frequency ratio for a rounding strategy $\placeholder$  in that equation, one can trivially see that the following relationships hold:
  \begin{equation}
  \alpha_k^{\floor{}} \leq \alpha \leq  \alpha_k^{\ceil{}}
  \qquad \text{and} \qquad
  \alpha_k^{\floor{}} \leq \alpha_k^{\round{}} \leq  \alpha_k^{\ceil{}}
  \qquad
  \forall k \in \NaturalNumbersWithoutZero
  \text{.}
  \end{equation}
  This means that rounding using the \mcode{floor} operator favors to strictly attain the specified frequency bin density at the cost of a few more excited lines (and hence less allottable power per line).
  The \mcode{ceil} function produces a grid that has more power per line at the cost of a reduced bin density.
  On the other hand, using \mcode{round} as in \eqref{eq:excitation:qlog:roundAlpha}, provides a grid that is somewhere in-between both extremes.
  For this reason, the grid constructed using $\alpha_k^{\round{}}$, i.e. by using the normal \mcode{round} function, has been used in the remaind of this chapter.
  \end{remark}

@@

  The arguments for a logarithmically spaced grid with an infinite span remain approximately valid for quasi-logarithmically spaced grids with a finite frequency support if the resonance peak of the system under test lies within the `bulk' of the frequency grid with finite support.
  \TODO{rule of thumb: when is it bulk or not? $f(\alpha,f_0,\Delta f, \wn, \damping)$}
  This can be understood intuitively by inspecting the derivatives that constitute the Fisher information matrix (see \figref{fig:excitation:derivatives}): those derivatives are significantly larger near the resonance frequency than at other frequencies.

  I.e. there should be a few excited frequency lines that are logarithmically spaced, outside of the resonance peak of the actual system.
  In practice, this means that systems with their resonance peak at the lowest and the very highest edge of the frequency band may not be observed with identical quality as systems at intermediate frequencies.

@@

  At the higher frequency bands, an approximately constant quality is still obtained for the (compensated) quasi-logarithmic signals.
  At lower frequencies, however, the logarithmically spaced signals perform only slightly better than the linearly spaced excitation signal.
  This is due to the fact that as the damping $\damping$ decreases, the $3\unit{dB}$ bandwidth of the resonance decreases and therefore less points can be fitted in this bandwidth at the lower frequencies.
  In our case, only $1$ line was excited in the $3 \unit{dB}$ bandwidth (respectively $0.4 \unit{Hz}$ and $0.8\unit{Hz}$ wide) of the systems with the lowest resonance frequencies.
  It is clearly unfeasible to estimate the five parameters in
  \eqref{eq:excitation:secondOrderSystemDTidentifiable} using this little information in the $3\unit{dB}$ bandwidth of the system.
  In other words, the resonances of the systems at low frequencies, can no longer be considered to lie within the bulk of the logarithmic part of the quasi-logarithmic frequency grids.
  We do see, however, that the variance in the vicinity of the resonance is lower than in the case of a linearly spaced excitation.
  \TODO{determine rule-of-thumb: resonance peak $\tuple{\wn,\damping}$ in bulk of grid $\tuple{\alpha,\omega_0,\Delta f}$}

@@

>>>>

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example} \label{eg:excitation:logarithmicGrid:matlab}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

  \begin{definition} \label{def:excitation:bulk}
  We denote the `\term{bulk}' of the quasi-logarithmic frequency grid as the frequency region where $\frac{f_k}{f_{k-1}} \approx \alpha$ can be attained.
  A necessary condition for $f_k$ to be part of the bulk is that $\Delta f \leq \left(\alpha -1 \right) f_{k-1}$.
  Equivalently, the bulk can be seen as the frequency interval $\RightOpenInterval{\frac{\Delta f}{\alpha - 1} , +\infty}$.
  \end{definition}

@@

  \begin{definition}
  For the implementation of a quasi-logarithmic frequency grid,  the approximate equality \eqref{eq:excitation:qlogGrid} does not define the grid unambiguously.
  The quasi-logarithmic grids depicted in this chapter are constructed using the relationship
  \begin{align}
       \label{eq:excitation:qlog:gridDef}
    f_k & = \alpha_k(n_k) f_{k-1}  
        \quad
            \forall k \in \set{1,\ldots,F}
\end{align}
with
\begin{align}
     \label{eq:excitation:qlog:roundAlpha}
      \alpha_k(N_k) & \isdef 
          \round{   \frac{\alpha^{n_k} f_{k-1}}{\Delta f}}    
           \frac{\Delta f}{f_{k-1}}
      \approx \alpha \text{, and}\\
         \label{eq:excitation:qlog:alphaPowerDef}
          n_k & \isdef 
            \min 
            \set{n | n \in \NaturalNumbers, \alpha_k(n) > 1}
     \text{.}
  \end{align}
  \end{definition}
  In the equation above, $\round{\placeholder}$ denotes rounding towards the nearest integer according to the \IEEEfloat standards and as implemented in the \MATLAB function \mcode{round}.
  This satisfies the approximate definition.

  \begin{remark}
  The use of different rounding strategies --- e.g. the \mcode{ceil} function $\ceil{}$  and \mcode{floor} function $\floor{}$ --- in \eqref{eq:excitation:qlog:roundAlpha}, allows for slightly different guarantees on the effective $\alpha_k$ attained by the grid.
  By denoting $\alpha_k^{\placeholder}$ as the effective frequency ratio for a rounding strategy $\placeholder$  in that equation, one can trivially see that the following relationships hold:
  \begin{equation}
  \alpha_k^{\floor{}} \leq \alpha \leq  \alpha_k^{\ceil{}}
  \qquad \text{and} \qquad
  \alpha_k^{\floor{}} \leq \alpha_k^{\round{}} \leq  \alpha_k^{\ceil{}}
  \qquad
  \forall k \in \NaturalNumbersWithoutZero
  \text{.}
  \end{equation}
  This means that rounding using the \mcode{floor} operator favors to strictly attain the specified frequency bin density at the cost of a few more excited lines (and hence less allottable power per line).
  The \mcode{ceil} function produces a grid that has more power per line at the cost of a reduced bin density.
  On the other hand, using \mcode{round} as in \eqref{eq:excitation:qlog:roundAlpha}, provides a grid that is somewhere in-between both extremes.
  For this reason, the grid constructed using $\alpha_k^{\round{}}$, i.e. by using the normal \mcode{round} function, has been used in the remainder of this chapter.
  \end{remark}

  \begin{example} \label{eg:excitation:quasilogarithmicGrid:matlab}
  An easy way to produce a quasi-logarithmic grid in \MATLAB is to first produce the logarithmic grid \code{fLog} as in  \egref{eg:excitation:logarithmicGrid:matlab} and reducing it to a quasi-logarithmic grid with resolution $\Delta f$ using \mcode[mathescape]{f = unique(round(fLog/$\Delta f$))*$\Delta f$}.
  Note that this is a computationally efficient way since both \mcode{logspace} and \mcode{unique} require at most $\bigO{ F }$ time, where \mcode[mathescape]{$F\;$ = numel(fLog)}.
  \end{example}

@@

  The arguments for a logarithmically spaced grid with an infinite span remain approximately valid for quasi-logarithmically spaced grids with a finite frequency support if the resonance peak of the system under test lies well within the `bulk' of the frequency grid with finite support, i.e. $\RightOpenInterval{\frac{2\pi \Delta f}{\alpha - 1}, +\infty}$.
  In particular for lightly-damped sub-systems, $\wdB \approx \ClosedInterval{(1-\damping)\wn, (1+\damping)\wn}$, such that the resonance peak is in the bulk of the frequency grid when
  \begin{equation} 
    \frac{2 \pi \Delta f}{alpha -1} \leq \left( 1 - \damping \right) \wn
    \text{.}
    \label{eq:excitation:systemInBulk:generic}
  \end{equation}
  For more reliable results, however, it is adviced to also have a few logarithmically spaced frequency lines, outside of the resonance peak of the actual system for a more robust result.
  In practice, this means that systems with their resonance peak at the lowest and the very highest edge of the frequency band may not be observed with identical quality as systems at intermediate frequencies.

  The validity of this approximation can be understood intuitively by inspecting the derivatives that constitute the Fisher information matrix (see \figref{fig:excitation:derivatives}): those derivatives are significantly larger near the resonance frequency than at other frequencies.

@@

  At the higher frequency bands, an approximately constant quality is still obtained for the (compensated) quasi-logarithmic signals.
  At lower frequencies, however, the logarithmically spaced signals perform only slightly better than the linearly spaced excitation signal.
  This is due to the fact that as the damping $\damping$ decreases, the $3\unit{dB}$ bandwidth of the resonance decreases and therefore less points can be fitted in this bandwidth at the lower frequencies.
  In our case, only $1$ line was excited in the $3 \unit{dB}$ bandwidth (respectively $0.4 \unit{Hz}$ and $0.8\unit{Hz}$ wide) of the systems with the lowest resonance frequencies.
  It is clearly unfeasible to estimate the five parameters in
  \eqref{eq:excitation:secondOrderSystemDTidentifiable} using this little information in the $3\unit{dB}$ bandwidth of the system.
  In other words, the resonances of the systems at low frequencies, can no longer be considered to lie within the bulk of the logarithmic part of the quasi-logarithmic frequency grids.
  This can also be understood from equation \eqref{eq:excitation:systemInBulk:generic}.
  In that equation, $\alpha = \alpha^{\star} = 1 + \damping$ such that
  \begin{equation}
    2 \pi \Delta f \leq (\damping - \damping^2) \wn
    \label{eq:excitation:systemInBulk:tuned}
  \end{equation}
  is the necessary condition for the system bandwidth to lie within the bulk of the frequency grid when the frequency spacing has been tuned according to \secref{sec:excitation:optimAlpha}.
  For this simulation, with $\Delta f=1\unit{Hz}$, $\damping=0.05$, this evaluates to $\wn \geq 132 \unit{rad/s} = 21 \unit{Hz}$.
  The system with $\wR = 20 \unit{Hz} < 21 \unit{Hz}$ is still well-identified, even though `only' $3$ lines are excited in its bandwidth.
  In that particular case, the frequency grid and the resonance peak are well-aligned ($\beta \approx 0$), which leads to a reduced variance (see also \figref{fig:excitation:worstCaseBeta}).

  We do see, however, that the variance in the vicinity of the resonance of every system is lower than in the case of a linearly spaced excitation

@@

\TODO{vuistregels herhalen in kader: $\alpha \sim \damping[\min]$, $\wn[\min] \sim \Delta \omega, \damping[\min]$}
\TODO{evt. plot vuistregels voor freq.grid.}
